---
title: "03_3di_Extraction"
output: html_document
---
# Libraries
```{r}
library(ape)
library(msa)
library(seqinr) # read.fasta
library(BioGeoBEARS) # sourceall
library(rvest) # for html_table()
library(phytools) # for midpoint.root
library(openxlsx) # for read.xlsx
```

# Sourcing
```{r}
sourceall("~/GitHub/str2phy/Rsrc/")
source("~/GitHub/str2phy/Rsrc/protein_bioinf_v1.R")
source("~/GitHub/str2phy/Rsrc/str2phy_v1.R")
sourceall("~/GitHub/str2phy/Rsrc/")
source("Script/Source.R")
```

```{bash}
cp ./Data/02_Distant_Homolog_Extraction/homologs_downloaded_renamed.fasta ./Data/03_3di_Character_Extraction/homologs_downloaded_renamed.fasta
```



# Running Chimera to get 3D structure of all amino acids from alphafold using Chimera

Alphafold returns cif file for each proteins Returns 3 files - \_alphafold.cif, \_alphafold.html, \_alphafold_table.html and txt file (defined in 'outdfs_fn')
```{r}
setwd("Data/03_3di_Character_Extraction")
fasta_fn = "homologs_downloaded_renamed.fasta" ## using the manual Jackhmmer hits  (along with the long branches)
tmpseqs = ape::read.FASTA(fasta_fn, type="AA")
numseqs = length(tmpseqs)
lines_list <- as.list(readLines("completed_ids.txt"))
lines_list <- lines_list[lines_list != ""]
outdfs = NULL; i=1
exclude_is = lines_list
for (i in 1:numseqs)
	{
	txt = paste0("Getting alphafold structure for sequence #", i, "/", numseqs)
	cat("\n")
	cat(txt)
	cat("\n")
	
	if ((i %in% exclude_is) == TRUE)
		{
		seqs = ape::read.FASTA(fasta_fn, type="AA")
		header = names(seqs)[[i]]
		gid = firstword(header)
		
		txt = paste0("Sequence #", i, "/", numseqs, ", aka '", gid, "', skipped as it is known to cause a problem.")
		cat("\n")
		cat(txt)
		cat("\n")
		outdf = rep(NA, times=13)
		names(outdf) = c("seqnum", "AlphaFold Model", "Query Sequence", "Identity %", "Coverage %", "cif_fn", "html_fn", "table_fn", "dump_fn", "chain_names_fn", "di3_fn", "aa_fn", "both_fn")
		outdfs = rbind(outdfs, outdf)
		next()
		} # END if (i %in% exclude_is)
	
	# Just get the CIF files
	print("same script")
	outdf = fasta_to_alphafold_cifs_modified(fasta_fn=fasta_fn, seqnum=i)
	print(outdf[1:5])
	outdfs = rbind(outdfs, outdf)
	
	outdfs_fn = "alphafolds.txt"
	write.table(x=outdfs, file=outdfs_fn, quote=FALSE, sep="\t", row.names=FALSE, col.names=TRUE)
	completed_id = names(tmpseqs[i])
	write(completed_id, file = "completed_ids.txt", append = TRUE, sep = "\n")
	
	} # END for (i in 1:numseqs)

## Returned Alphafold structure for only 377 sequences
```

# Extracting the html info - Alphafolds.txt was modified manually to exclude the lines without .cif file

```{r}
setwd("Data/03_3di_Character_Extraction")
source("~/GitHub/str2phy/Rsrc/str2phy_v1.R")
outdfs2_fn = "alphafolds.txt"
outdfs = read.table(outdfs2_fn, quote="", sep="\t", header=TRUE, na.strings = c("NA", ""))
outdfs
names_outdfs = c("seqnum", "AlphaFold Model", "Query Sequence", "Identity %", "Coverage %", "cif_fn", "html_fn", "table_fn", "dump_fn", "chain_names_fn", "di3_fn", "aa_fn", "both_fn")
names(outdfs) = names_outdfs
outdfs
alphafold_dfs = NULL

for (i in 1:376)
	{
	txt = paste0("Reading ChimeraX --alphafold table returned for sequence #", i, "/", numseqs)
	cat("\n")
	cat(txt)
	
	html_fn = outdfs$html_fn[i]
	print(html_fn)
	table_fn = outdfs$table_fn[i]
	print(table_fn)
	
	alphafold_df = alphafold_html_to_df(html_fn, table_fn)
	alphafold_df
	
	alphafold_dfs = rbind(alphafold_dfs, alphafold_df)
	}
cat("\n...done\n")
alphafold_dfs
dim(outdfs)

# Write over, if needed
outdfs2 = outdfs
outdfs2[,"AlphaFold Model"] = alphafold_dfs[,"AlphaFold Model"]
outdfs2[,"Query Sequence"] = alphafold_dfs[,"Query Sequence"]
outdfs2[,"Identity %"] = alphafold_dfs[,"Identity %"]
outdfs2[,"Coverage %"] = alphafold_dfs[,"Coverage %"]

outdfs2_fn = "alphafolds.txt"
write.table(x=outdfs2, file=outdfs2_fn, quote=FALSE, sep="\t", row.names=FALSE, col.names=TRUE)

# Look at the output table
View(outdfs2)
```

# Concatenate all _3di.fasta files into a big one
```{bash}

cat ./Data/03_3di_Character_Extraction/*_alphafold_3di.fasta > ./Data/03_3di_Character_Extraction/3di_unaligned.fasta
```

# Concatenate all aa.fasta files into a big one
```{bash}

cat ./Data/03_3di_Character_Extraction/*_alphafold_aa.fasta > ./Data/03_3di_Character_Extraction/aa_unaligned.fasta
```

# Renaming to remove .cif from the sequence names
```{bash}
sed -i -e 's/_alphafold.cif//g' ./Data/03_3di_Character_Extraction/3di_unaligned.fasta
sed -i -e 's/_alphafold.cif//g' ./Data/03_3di_Character_Extraction/aa_unaligned.fasta
```

## Extracting only the sequennce required from previous analysis file
```{r}
xlsfn = "Data/RenamingSequences.xlsx"
xls = openxlsx::read.xlsx(xlsxFile=xlsfn, sheet=1, startRow=1, colNames=TRUE)

seqids = as.character(xls$AccessionNumber)
new_names <- as.character(xls$protein)

lookup <- setNames(new_names, seqids)

tmpfn_aa = "Data/03_3di_Character_Extraction/aa_unaligned.fasta"
tmpfn_3di <- "Data/03_3di_Character_Extraction/3di_unaligned.fasta"
aa_fasta <- read.fasta (file = "Data/03_3di_Character_Extraction/aa_unaligned_PreviousAnalysis.fasta")
di3_fasta <- read.fasta (file = "Data/03_3di_Character_Extraction/3di_unaligned_PreviousAnalysis.fasta")
filtered_seqs <- list()
filtered_names <- c()

for (i in seq_along(aa_fasta)) {
  # Get the sequence name
  seq_name <- attr(aa_fasta[[i]], "name")
  
  # Extract the part before the first hyphen
  acc <- strsplit(seq_name, "_")[[1]][1]
  # Check if it is in seqids
  if (acc %in% seqids) {
    new_name <- lookup[[acc]]
    filtered_seqs[[seq_name]] <- aa_fasta[[i]]
    filtered_names <- c(filtered_names, new_name)
  }
}
length(filtered_seqs)
write.fasta(sequences = filtered_seqs, names = filtered_names, file.out = tmpfn_aa)

filtered_seqs_3di <- list()
filtered_names_3di <- c()
for (i in seq_along(di3_fasta)) {
  # Get the sequence name
  seq_name <- attr(di3_fasta[[i]], "name")
  
  # Extract the part before the first hyphen
  acc <- strsplit(seq_name, "_")[[1]][1]
  
  # Check if it is in seqids
  if (acc %in% seqids) {
    new_name <- lookup[[acc]]
    filtered_seqs_3di[[seq_name]] <- di3_fasta[[i]]
    filtered_names_3di <- c(filtered_names, new_name)
  }
}
length(filtered_seqs_3di)
write.fasta(sequences = filtered_seqs_3di, names = filtered_names_3di, file.out = tmpfn_3di)
```

